---
title: "Reproducible Research: Course Project 2"
output: 
  html_document:
    keep_md: true
---

####Introduction

Storms and other severe weather events can cause both public health and economic problems for communities and municipalities. Many severe events can result in fatalities, injuries, and property damage, and preventing such outcomes to the extent possible is a key concern.

This project involves exploring the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database. This database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.


###Data Download
Unzip and Load the from bzp2

```{r}
df <- read.csv("data/repdata_data_StormData.csv.bz2")
```

###Analyzing Data
First of all, evaluate the total fatalities and total injuries for each event type help us to understand the health impact.

```{r}
library(dplyr)

df.fatalities <- arrange(
    summarise(
        group_by(
            select(
                df,
                EVTYPE,
                FATALITIES
            ),
            EVTYPE
        ),
        total.fatalities = sum(FATALITIES)
    ),
    -total.fatalities
)

head(df.fatalities, 10)
```

And the injuries ...

```{r}
df.injuries <- arrange(
    summarise(
        group_by(
            select(
                df,
                EVTYPE,
                INJURIES
            ),
            EVTYPE
        ),
        total.injuries = sum(INJURIES)
    ),
    -total.injuries
)

head(df.injuries, 10)

```

####Study the Economical Impact
There`s two economic impacts:  
[1] Property Damage(PROPDMG)
[2] Crop Damage(CROPDMG)

The actual damage in $USD is indicated by PROPDMGEXP and CROPDMGEXP parameters.

```{r}
# Define the dataframe
df.damage <- select(df, EVTYPE, PROPDMG,PROPDMGEXP,CROPDMG,CROPDMGEXP)

# Ordered by PROPDMGEXP
Symbol <- sort(unique(as.character(df.damage$PROPDMGEXP)))
# List of multipliers
Multiplier <- c(0,0,0,1,10,10,10,10,10,10,10,10,10,10^9,10^2,10^2,10^3,10^6,10^6)

convert.Multiplier <- data.frame(Symbol, Multiplier)

df.damage$Prop.Multiplier <- convert.Multiplier$Multiplier[match(df.damage$PROPDMGEXP, convert.Multiplier$Symbol)]

df.damage$Crop.Multiplier <- convert.Multiplier$Multiplier[match(df.damage$CROPDMGEXP, convert.Multiplier$Symbol)]

df.damage <- df.damage %>% mutate(PROPDMG = PROPDMG*Prop.Multiplier) %>% mutate(CROPDMG = CROPDMG*Crop.Multiplier) %>% mutate(TOTAL.DMG = PROPDMG+CROPDMG)

df.damage.total <- df.damage %>% group_by(EVTYPE) %>% summarize(TOTAL.DMG.EVTYPE = sum(TOTAL.DMG))%>% arrange(-TOTAL.DMG.EVTYPE) 

head(df.damage.total,10)
```

###Health Impact

```{r}
library(ggplot2)

g <- ggplot(
    df.fatalities[1:10,],
    aes(x=reorder(EVTYPE, -total.fatalities), y=total.fatalities))+
    geom_bar(stat="identity")+
    theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=1))+
    ggtitle("Top 10 Events with Highest Total Fatalities")+
    labs(x="EVENT TYPE", y="Total Fatalities")

g
```


```{r}
g <- ggplot(
    df.injuries[1:10,],
    aes(x=reorder(EVTYPE, -total.injuries), y=total.injuries))+
    geom_bar(stat="identity")+
    theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=1))+
    ggtitle("Top 10 Events with Highest Total Injuries")+
    labs(x="EVENT TYPE", y="Total Injuries")

g
```

###Economic Impact

```{r}
g <- ggplot(
    df.damage.total[1:10,],
    aes(x=reorder(EVTYPE, -TOTAL.DMG.EVTYPE), y=TOTAL.DMG.EVTYPE))+
    geom_bar(stat="identity")+
    theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=1))+
    ggtitle("Top 10 Events with Highest Economic Impact")+
    labs(x="EVENT TYPE", y="Total Economic Impact ($USD)")

g
```

